---
title: "Resistance Is Futile: Anti-Trump Demonstrations and their Effect on Legislation"
author: "Mac Bagwell"
date: "2020-03-11"
output: 
  github_document:
    toc: true
---

```{r message=FALSE}
library(tidyverse)
library(lubridate)
library(modelr)

START <- "2017-02-02"
END <- "2018-12-31"

LAMBDA <- 0.95
WINDOW <- 90

file_protests_population <- here::here("data/protests_population.rds")
file_member_averages <- here::here("data/member_averages.rds")
file_votes <- here::here("data/votes.rds")
file_class <- here::here("data/senate-class.rds")

#===============================================================================

# Code
ewma <- function(x, lambda, init = median(x, na.rm = TRUE)) {
	p <- init
  if_else(!is.finite(x), 0, x) %>% 
	  map_dbl(~ (p <<- p * lambda + (.) * (1 - lambda)))
}

# Aggregate data
protests_population <- 
  file_protests_population %>% 
  read_rds() %>% 
  filter(pro_anti_trump == "anti")

member_averages <- 
  file_member_averages %>% 
  read_rds() %>% 
  filter(congress == 115, party != "Independent", votes > 75) %>% 
  left_join(
    file_class %>% 
      read_rds(),
    by = c("last_name", "state")
  )

votes <- 
  file_votes %>% 
  read_rds() 

# Time series data
state_protests_ts <- 
  tibble(date = seq(ymd(START), ymd(END), by = "1 day")) %>% 
  left_join(
    protests_population %>% 
      count(
        state, 
        date = floor_date(date, unit = "day"), 
        wt =  attendance / (population / 1e5),
        name = "protest_activity"
      ) %>% 
      pivot_wider(
        names_from = state,
        values_from = protest_activity,
        values_fill = list(protest_activity = 0L)
      ),
    by = "date"
  ) %>% 
  mutate_at(
    vars(-date),
    ewma,
    lambda = LAMBDA
  ) %>%
  pivot_longer(
    cols = -date,
    names_to = "state",
    values_to = "protest_attendance"
  )

trump_score_ts <- 
  tibble(date = seq(ymd(START), ymd(END), by = "1 day")) %>% 
  left_join(
    votes %>% 
      semi_join(
        member_averages %>% 
          filter(congress == 115L, votes >= 80, party != "Independent"),
        by = "bioguide"
      ) %>% 
      mutate(date = as_date(voted_at)) %>% 
      group_by(date, bioguide, last_name, state, party) %>% 
      summarize(
        agrees = sum(agree),
        votes = n()
      ) %>% 
      pivot_wider(
        names_from = c(bioguide, last_name, state, party),
        values_from = c(agrees, votes)
      ),
    by = "date"
  ) %>% 
  replace_na(list(agrees = 0L, votes = 0L)) %>% 
  mutate_at(
    vars(
      starts_with("agrees"), 
      starts_with("votes")
    ),
    list(
      forward = 
        ~ RcppRoll::roll_suml(., n = WINDOW, fill = NA_integer_, na.rm = TRUE),
      past = 
        ~ lag(.) %>% 
          RcppRoll::roll_sumr(., n = WINDOW, fill = NA_integer_, na.rm = TRUE)
    )
  ) %>% 
  select(date, ends_with("forward"), ends_with("past")) %>% 
  pivot_longer(
    cols = -date,
    names_to = c("variable", "bioguide", "last_name", "state", "party", "window"),
    names_sep = "_",
    values_to = "value"
  ) %>% 
  pivot_wider(
    names_from = c(variable, window),
    values_from = value
  ) %>%
  mutate(
    trump_score_forward = agrees_forward / votes_forward,
    trump_score_past = agrees_past / votes_past,
    trump_score_diff = trump_score_forward - trump_score_past,
  ) %>% 
  group_by(date, party) %>% 
  mutate(trump_score_party_index = percent_rank(trump_score_diff))
```

## Introduction

The first two years of the Trump presidency was marked by extremely high levels
of political activism. According the the Crowd Counting Consortium, 
from February 2017 to December 2018, there were more than 16,500 distinct protests
with a total attendance of over 7.5 million people, which is more than 2% of the
US population. Most of these protests were aimed at Donald Trump's agenda.
Nearly 55% of all protests in the US, accounting for 72% of total protest attendance,
had an anti-Trump message in the years 2017 and 2018. 

The theory of the case for this explosion of protests is 
typically that they work to galvanize Democrats' legislative resistance 
to the president's agenda, particularly in the Senate. [^1] 
Some in academia have concluded that public demonstrations have succeeded
in doing this. [^2] We are going to put that theory to the test ourselves.

## Data

### Trump Score

FiveThirtyEight tracks a figure that they call "Trump Score". Trump Score is defined
for a congressperson as the percentage of all that senator's roll-call votes
that are in favor of a bill that Trump supports or against a bill that he opposes. 
Naturally, we expect this number to be lower among legislators exhibiting
increased resistance to the Trump agenda. We will use this as a measure of
this resistance. For our purposes, we will focus only on senators during the 115th
Congress (2017-18) that did not end their term abnormally. For example, we exclude
Jeff Sessions from our analysis, who left the Senate to become Attorney General,
as well as the senators that took his place.

FiveThirtyEight also provides vote-level data for each roll-call vote used
to compute the Trump score. This makes it possible to track how Trump scores
evolve over time for each senator. 

Here is a sample of rows from the aggregated table:

```{r}
member_averages %>% 
  head() %>% 
  knitr::kable()
```

And from the detailed table:

```{r}
votes %>% 
  head() %>% 
  knitr::kable()
```


### Protests

The Crowd Counting Consortium collects and publishes data on protests that occured in the United
States since February 2017, sourced largely on from coverage. They label every protest with the date of the protest, 
state in which it happened, an estimate of the number of people in attendance [^3], 
and whether the protest was pro-Trump in messaging, anti-Trump in messaging, or neither.
For our analysis, we only consider anti-Trump protests with attendance estimates
present in the data, though it is worth noting that about a third of crowd estimates are missing across all
sentiments (see table below).

```{r echo=FALSE}
protests_population %>% 
  drop_na(pro_anti_trump) %>% 
  count(pro_anti_trump, missing = is.na(attendance)) %>% 
  group_by(pro_anti_trump) %>% 
  summarize(prop = (last(n) / sum(n) * 100) %>% round(2)) %>% 
  mutate(
    pro_anti_trump = 
      recode(
        pro_anti_trump,
        anti = "Anti-Trump",
        neutral = "Neutral",
        pro = "Pro-Trump"
      ),
    prop = str_c(prop, "%")
  ) %>% 
  knitr::kable(col.names = c("Sentiment", "% Missing Attendance Estimates"))
```

Here is a sample of rows from the relevant table:

```{r}
protests_population %>% 
  head(10) %>% 
  knitr::kable()
```

## Analysis

We take two different approaches to this analysis. 

* First, we examine Trump scores and protest activity by state on aggregate 
through the 115th Congress to determine if there are any relationships between the two
in the *long term*.
* Second, we examine changes in a windowed Trump score over time and a moving
average of protest activity by state to see if there are any point-in-time effects
in the *short term*.

### The Long Term

A natural thing to do if we want to discern a relationship between Trump Score
and protests would be to plot the Trump score of each senator against the intensity of protests in their
state, measured by total protest attendance per 100,000 people, in years 2017 and 2018.
We obtain the following:

```{r fig.asp=0.75, fig.width=10}
protests_population %>% 
  count(
    state, 
    name = "protest_activity", 
    wt = attendance / (population / 1e5)
  ) %>% 
  left_join(member_averages, by = "state") %>% 
  ggplot(aes(protest_activity, agree_pct, fill = party)) +
  geom_smooth(aes(color = party), method = "lm", formula = y ~ log(x)) +
  geom_point(shape = 21, color = "black") +
  scale_color_manual(
    breaks = c("Republican", "Democrat"),
    values = c(Republican = "red3", Democrat = "royalblue"),
    aesthetics = c("color", "fill")
  ) +
  scale_x_log10(
    breaks = scales::breaks_log(n = 10)
  ) +
  scale_y_continuous(
    labels = scales::label_percent()
  ) +
  coord_fixed(ratio = 1.5) +
  theme(legend.position = "bottom") +
  labs(
    title = "Senator Trump Score vs. State Protest Activity, 115th Congress",
    subtitle = "Independents omitted",
    caption = "Sources: FiveThirtyEight, Crowd Counting Consortium, ACS 2017",
    x = "Protest Attendance per 100,000 People, Feb. 2017 - Dec. 2018",
    y = "Trump Score",
    color = NULL,
    fill = NULL
  )
```

This certainly seems dispositive at first glance; we observe a pretty large gap between the Republicans
and Democrats, but controlling for party, it seems like protest attendance has a small but noticeable
effect on senators' Trump scores. However, this ignores an obvious latent variable that could correlate 
both with Trump score and protest attendance: Trump's favorability by state. 

The plots below that Trump's performance in each state has a sizeable impact both
on senators' Trump scores and on protest attendance.

```{r fig.asp=0.75, fig.width=10}
protest_trump_state <- 
  protests_population %>% 
  count(state, name = "protest_level", wt = attendance / (population / 1e5)) %>% 
  left_join(member_averages, by = "state")

protest_trump_state %>% 
  ggplot(aes(net_trump_vote, protest_level)) +
  geom_vline(xintercept = 0, color = "white", size = 1.5) +
  geom_smooth(method = "lm") +
  geom_text(aes(label = state)) +
  scale_x_continuous(labels = scales::label_percent(scale = 1)) +
  scale_y_log10(breaks = scales::breaks_log(n = 9)) +
  labs(
    title = "Protest Attendance per 100,000 people vs. Net Trump Vote",
    subtitle = "States that favored Trump are less active",
    caption = "Sources: FiveThirtyEight, Crowd Counting Consortium, ACS 2017",
    x = "Trump's Margin of Victory, 2016",
    y = "Protest Attendance per 100,000 people, Feb. 2017 - Dec. 2018"
  )
```

```{r fig.asp=0.75, fig.width=10}
member_averages %>% 
  ggplot(aes(net_trump_vote, agree_pct)) +
  geom_hline(yintercept = 0.5, size = 1.5, color = "white") +
  geom_vline(xintercept = 0, size = 1.5, color = "white") +
  geom_smooth(aes(color = party), method = "lm") +
  geom_point(aes(fill = party), shape = 21, color = "black") +
  scale_x_continuous(
    breaks = scales::breaks_width(10),
    labels = scales::label_percent(scale = 1)
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(0.1),
    labels = scales::label_percent(accuracy = 1)
  ) +
  scale_fill_manual(
    breaks = c("Republican", "Democrat"),
    values = c("red3", "royalblue"),
    aesthetics = c("color", "fill")
  ) +
  coord_fixed(ratio = 60) +
  theme(legend.position = "bottom") +
  labs(
    title = "Trump Score vs. State Net Trump Vote",
    subtitle = "States that favored Trump have senators that vote with Trump more often",
    caption = "Sources: FiveThirtyEight",
    x = "State Net Trump Vote (2016)",
    y = "Trump Score",
    color = NULL,
    fill = NULL
  )
```

This stands to obfuscate the true effect of protest activity on senators' Trump
scores; it might be the case that protest attendance simply serves as a proxy for
the only real underlying motivator of Trump scores which is Trump's favorability
by state. To control for this, we took the residuals from the linear models shown
in the previous two plots given by

`trump_score = a_0 + a_1 . net_trump_vote + a_2 . party + a_3 . trump_vote . party`

`protest_attendance = a_0 + a_1 . net_trump_vote`

and plotted them against each other.

```{r fig.asp=0.75, fig.width=10}
protest_trump_mod <- 
  protest_trump_state %>% 
  fit_with(
    lm, 
    list(
      protest = log(protest_level) ~ net_trump_vote,
      trump = agree_pct ~ net_trump_vote * party
    )
  )

protest_trump_state %>% 
  spread_residuals(!!! protest_trump_mod) %>% 
  ggplot(aes(protest, trump, fill = party)) +
  geom_hline(yintercept = 0, size = 1.5, color = "white") +
  geom_vline(xintercept = 0, size = 1.5, color = "white") +
  geom_point(shape = 21, size = 2, color = "black") +
  scale_fill_manual(
    breaks = c("Republican", "Democrat"),
    values = c(Republican = "red3", Democrat = "royalblue")
  ) +
  scale_y_continuous(
    labels = scales::label_percent()
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = "Residual Trump Score vs. Residual Protest Activity during the 115th Congress",
    subtitle = "Distribution does not suggest residual effects",
    caption = "Variables controlled for Trump's 2016 margin of victory by state\nSources: FiveThirtyEight, Crowd Counting Consortium",
    x = "Residual Log(Protest Activity), Feb. 2017 - Dec. 2018",
    y = "Residual Trump Score",
    fill = NULL
  ) 
```

It is clear that no discernible trend exists among these two variables. This strongly
suggests that protest activity does not really move the needle that much on a senators'
resistance to Trump over long periods of time. 

### The Short Term

While the above analysis shows pretty definitively that there
isn't much to be gained in the long run from public demonstrations, it might
be the case that bursts of protests temporarily affect senators' voting patterns.

To test this model, we need to come up with point-in-time versions of the
protest attendance and Trump score signals, which will allow us to see how these
correlate over time.

To convert Trump score into a time-series statistic, we calculate at every day
for every senator the Trump score of that senator using the votes over the next
90 days. Below are plots of the distribution of this signal.

```{r fig.asp=0.5, fig.width=10}
trump_score_ts %>% 
  drop_na(trump_score_diff) %>% 
  group_by(date) %>% 
  summarize(
    pct_25 = quantile(trump_score_diff, 0.25, na.rm = TRUE),
    pct_50 = quantile(trump_score_diff, 0.5, na.rm = TRUE),
    pct_75 = quantile(trump_score_diff, 0.75, na.rm = TRUE)
  ) %>% 
  ggplot(aes(date)) +
  geom_ribbon(aes(ymin = pct_25, ymax = pct_75), alpha = 0.25) +
  geom_line(aes(y = pct_50), size = 1) +
  geom_hline(yintercept = 0) +
  scale_x_date(
    date_breaks = "2 months",
    date_labels = "%b %y"
  ) +
  scale_y_continuous( 
    breaks = scales::breaks_width(0.1),
    labels = scales::label_percent(accuracy = 1)
  ) +
  scale_fill_manual(
    breaks = c("Republican", "Democrat"),
    values = c(Republican = "red", Democrat = "royalblue"),
    aesthetics = c("color", "fill")
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  ) +
  labs(
    title = str_glue("Middle 50% of {WINDOW}D Trump Score Change, 115th Congress"),
    caption = "Source: FiveThirtyEight",
    x = NULL,
    y = "Trump Score",
    fill = NULL,
    color = NULL
  )

trump_score_ts %>% 
  drop_na(trump_score_diff) %>% 
  group_by(date, party) %>% 
  summarize(
    pct_25 = quantile(trump_score_diff, 0.25, na.rm = TRUE),
    pct_50 = quantile(trump_score_diff, 0.5, na.rm = TRUE),
    pct_75 = quantile(trump_score_diff, 0.75, na.rm = TRUE)
  ) %>% 
  ggplot(aes(date, color = party, fill = party)) +
  geom_hline(yintercept = 0, color = "white", size = 1.5) +
  geom_ribbon(aes(ymin = pct_25, ymax = pct_75), alpha = 0.25) +
  geom_line(aes(y = pct_50, group = party), size = 1) +
  scale_x_date(
    date_breaks = "2 months",
    date_labels = "%b %y"
  ) +
  scale_y_continuous( 
    breaks = scales::breaks_width(0.1),
    labels = scales::label_percent(accuracy = 1)
  ) +
  scale_fill_manual(
    breaks = c("Republican", "Democrat"),
    values = c(Republican = "red3", Democrat = "royalblue"),
    aesthetics = c("color", "fill")
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  ) +
  labs(
    title = str_glue("Middle 50% of {WINDOW}D Trump Score Change by Party, 115th Congress"),
    caption = "Source: FiveThirtyEight",
    x = NULL,
    y = "Trump Score",
    fill = NULL,
    color = NULL
  )
```


To convert protest attendance into a time-series signal, we take
an exponential moving average of daily protest attendance per 100,000 people. [^4]
To see an illustration of how this moving average works, below we show the daily national
protest attendance and its exponential moving average.

```{r fig.asp=0.5, fig.width=10}
protests_population %>% 
  count(date, wt = attendance, name = "attendance") %>% 
  mutate(ewma_attendance = ewma(attendance, lambda = LAMBDA)) %>% 
  ggplot(aes(date)) +
  geom_segment(aes(xend = date, y = 1, yend = 1 + attendance), alpha = 0.5) +
  geom_line(aes(y = 1 + ewma_attendance), color = "red") +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y") +
  scale_y_log10(n.breaks = 6, labels = scales::label_number()) + 
  labs(
    title = "National Protest Attendance over Time",
    subtitle = "Red line represents exponential moving average",
    caption = "Source: Crowd Counting Consortium",
    x = NULL,
    y = "National Protest Attendance per 100,000"
  )
```


The senator-level statistic we will analyze will be the spearman correlation coefficient of 
the following two day-level variables

* The difference between each senator's Trump score over the past 90 days and their
Trump score over the next 90 days.
* The log of the exponential moving average of
daily protest attendance in each senator's state. [^5]

Below is the distribution of the correlation coefficients of all senators.

```{r fig.asp=0.5, fig.width=10}
corrs <- 
  state_protests_ts %>% 
  left_join(
    trump_score_ts %>% 
      semi_join(member_averages, by = "bioguide"), 
    by = c("date", "state")
  ) %>% 
  group_by(bioguide) %>% 
  summarize(
    r_diff = 
      cor(
        log(protest_attendance), 
        trump_score_diff, 
        method = "spearman", 
        use = "pair"
      )
  ) %>% 
  left_join(member_averages, by = "bioguide")

corrs %>% 
  ggplot(aes(r_diff, fill = party)) +
  geom_vline(xintercept = 0, color = "grey50", size = 1.5) +
  geom_dotplot(binwidth = 0.02, alpha = 0.8) +
  scale_x_continuous(
    breaks = scales::breaks_width(0.1)
  ) +
  scale_fill_manual(
    breaks = c("Democrat", "Republican"),
    values = c(Democrat = "royalblue", Republican = "red3")
  ) + 
  theme(legend.position = "bottom") +
  labs(
    title = str_glue("Correlation {WINDOW}D Trump Score Change and Protest Activity"),
    subtitle = "Distribution suggests legislators compromise more often following increased protest activity",
    caption = "Spearman method used to compute correlation\nSources: FiveThirtyEight, Crowd Counting Consortium",
    x = "Correlation Coefficient",
    y = NULL,
    fill = NULL
  )

corrs %>% 
  ggplot(aes(party, r_diff, fill = party)) +
  geom_hline(yintercept = 0, color = "grey50", size = 1.5) +
  geom_boxplot(alpha = 0.8, show.legend = FALSE) +
  scale_y_continuous(
    breaks = scales::breaks_width(0.1)
  ) +
  scale_fill_manual(
    breaks = c("Democrat", "Republican"),
    values = c(Democrat = "royalblue", Republican = "red3")
  ) + 
  coord_flip() +
  theme(legend.position = "bottom") +
  labs(
    title = str_glue("Correlation of {WINDOW}D Trump Score Change and Protest Activity"),
    subtitle = "Distribution suggests legislators compromise more often following increased protest activity",
    caption = "Spearman method used to compute correlation\nSources: FiveThirtyEight, Crowd Counting Consortium",
    x = NULL,
    y = "Correlation Coefficient",
    fill = NULL
  )
```

The distribution is pretty surprising. This suggests that legislators are actually
more likely to __compromise__ following increased intensity of public demonstrations.

To put it bluntly, I don't know why this happens. One explanation
I can think of is that protests could be coming just __after__ periods of intensified resistance
in Congress, which could in turn mean that the protests signal some kind of regression
to the mean moreso than they signal the period of increased resistance. It could
also mean that lawmakers are actually more interested in compromise in the name
of getting something done when they see that their constituents are unhappy. This is definitely
something worth investigating further, but might require more sophisticated data
or knowledge of lawmaking than is available to me right now.

In any case, I broke this coefficient out by by senatorial class. 

```{r fig.asp=0.75, fig.width=10}
party_medians <- 
  corrs %>% 
  group_by(party) %>% 
  summarize(r_diff = median(r_diff, na.rm = TRUE))

corrs %>% 
  ggplot(aes(r_diff, fill = party)) +
  geom_vline(xintercept = 0, color = "grey50", size = 1.5) +
  geom_vline(
    aes(color = party, xintercept = r_diff), 
    data = party_medians, 
    size = 1.5,
    show.legend = FALSE
  ) +
  geom_dotplot(binwidth = 0.02, alpha = 0.8) +
  geom_text(
    aes(color = party),
    data = party_medians,
    y = Inf,
    hjust = 1.1,
    vjust = 1.7,
    label = "Party Median",
    show.legend = FALSE
  ) +
  scale_x_continuous(
    breaks = scales::breaks_width(0.1)
  ) +
  scale_fill_manual(
    breaks = c("Democrat", "Republican"),
    values = c(Democrat = "royalblue", Republican = "red3"),
    aesthetics = c("color", "fill")
  ) + 
  facet_grid(
    rows = vars(class), 
    labeller = labeller(.rows = function(x) str_c("Class ", x))
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = str_glue("Correlation of {WINDOW}D Trump Score Change and Protest Activity by Senate Class"),
    subtitle = "Class 1 up for re-election in 2018, contains many lawmakers with high correlation coefficients",
    caption = "Spearman method used to compute correlation\nSources: FiveThirtyEight, Crowd Counting Consortium, Wikipedia",
    x = "Correlation Coefficient",
    y = NULL,
    fill = NULL
  ) 

corrs %>% 
  ggplot(aes(party, r_diff, fill = party)) +
  geom_hline(yintercept = 0, color = "grey50", size = 1.5) +
  geom_hline(
    aes(yintercept = r_diff, color = party), 
    data = party_medians, 
    size = 1.5,
    show.legend = FALSE
  ) +
  geom_boxplot(alpha = 0.8, show.legend = FALSE) +
  geom_text(
    aes(color = party),
    data = party_medians,
    x = Inf,
    hjust = 1.1,
    vjust = 1.7,
    label = "Party Median",
    show.legend = FALSE
  ) +
  scale_x_discrete(expand = c(0.5, 0.2)) +
  scale_y_continuous(
    breaks = scales::breaks_width(0.1)
  ) +
  scale_fill_manual(
    breaks = c("Democrat", "Republican"),
    values = c(Democrat = "royalblue", Republican = "red3"),
    aesthetics = c("color", "fill")
  ) + 
  coord_flip() +
  facet_grid(
    rows = vars(class), 
    labeller = labeller(.rows = function(x) str_c("Class ", x))
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = str_glue("Correlation of {WINDOW}D Trump Score Change and Protest Activity by Senate Class"),
    subtitle = "Class 1 up for re-election in 2018, contains many lawmakers with high correlation coefficients",
    caption = "Spearman method used to compute correlation\nSources: FiveThirtyEight, Crowd Counting Consortium, Wikipedia",
    x = NULL,
    y = "Correlation Coefficient",
    fill = NULL,
    color = NULL
  )
```

It is difficult to draw sweeping conclusions given the small number of senators in each class, 
but it seems like the senators in class 1, which was up for re-election in 2018, had a relatively
large number of senators who were highly likely to compromise following increased
protest activity. This could strengthen the argument that lawmakers prefer the
optics of compromise when faced with an upcoming election.

## Conclusion

Despite our best efforts, we found that in general, protest attendance was not
a very good indicator of senators' resistance to Trumpian policies. More than
anything, it is just a weak proxy for the general political attitudes of a state
which is a stronger indicator of Trump scores along with party affiliation.

In fact, on the contrary, it is possible that protests are a (fairly noisy) indicator that
lawmakers are more likely to seek compromise in the short-term as opposed
to a long-term resistance. It is unclear what if any causality exists here; it might
be a sign that senators up for re-election were seeking to project a particular image
in the name of winning their election. The evidence for this link is weak, though,
and requires more investigation in any case.

But if you're a citizen that feels passionately about stopping Donald Trump's policy agenda in general, 
you're probably better off waiting for election day and getting to the polls rather than
taking to the streets.

### Footnotes

[^1]: [Gambino and Gabbatt, 2017](https://www.theguardian.com/us-news/2017/feb/19/democrats-resistance-donald-trump-we-will-replace-you)

[^2]: [Miras, N.S. (2019), Polls and Elections: Resistance Is Not Futile: Anti‐Trump Protest and Senators’ Opposition to President Trump in the 115th Congress. Presidential Studies Quarterly, 49: 932-958. doi:10.1111/psq.12575](https://onlinelibrary.wiley.com/doi/full/10.1111/psq.12575)

[^3]: In some cases, they provide a single value as an estimate, and in others, they
provide a range. In the cases where they provide a range, we imputed the single-value
estimate with the geometric mean of the endpoints of the range.

[^4]: See [Wikipedia](https://en.wikipedia.org/wiki/Moving_average#Exponential_moving_average) for
details. We use a decay factor (α) of 0.05, though the exact value should not be of much
consequence for the purposes of our analysis.

[^5]: We use the spearman method instead of the usual pearson method because
it is more robust to outliers and non-linear relationships.
