---
title: "Trump Score"
author: "Mac Bagwell" 
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
= 
```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(lubridate)
# Parameters
START_115 <- ymd("2017-02-02")
END_115 <- ymd("2018-12-31")
votes_file <- here::here("c01-own/data/votes.rds")
member_averages_file <- here::here("c01-own/data/member_averages.rds")
window <- 180L
#===============================================================================
votes_df <- read_rds(votes_file)
member_averages_df <- read_rds(member_averages_file)
```

## Trump Scores Distribution

Unsurprisingly, Democrats and Republicans occupy distinct intervals
on the Trump Score scale. There is a bit more variance in Democrats'
Trump Score, though, with some senators agreeing with Trump more often than
not during the 115th congress.

```{r}
member_averages_df %>% 
  filter(congress == 115, party != "Independent") %>% 
  ggplot(aes(agree_pct, fill = party)) +
  geom_vline(xintercept = 0.5) +
  geom_dotplot(binwidth = 0.02) +
  scale_x_continuous(
    breaks = scales::breaks_width(0.1),
    labels = scales::label_percent(accuracy = 1)
  ) +
  scale_fill_manual(
    breaks = c("Republican", "Democrat"),
    values = c("red", "royalblue")
  ) +
  coord_cartesian(xlim = c(0.05, 1.05), expand = FALSE) +
  theme(
    axis.text.y = element_blank(),
    legend.position = "bottom"
  ) +
  labs(
    title = "Trump Score Distribution, 115th Congress",
    x = "Trump Score",
    y = "",
    fill = ""
  ) +
  ggsave("figures/trump_score_distribution.png")
```

## Trump Score vs. Trump Support

Much of the variation in Democrats' Trump Scores can be attributed to the
political tastes of their constituent state. The Democrats with outlier Trump
Scores hailed from states that Trump took by a veritable landslide in 2016.
This is not surprising, but it complicates our analysis of the effects of
protesting on senatorial votes, since presumably, protest activity will
correlate significantly with political tastes.

```{r}
member_averages_df %>% 
  filter(congress == 115, party != "Independent") %>% 
  ggplot(aes(net_trump_vote, agree_pct)) +
  geom_hline(yintercept = 0.5, size = 1.5, color = "white") +
  geom_vline(xintercept = 0, size = 1.5, color = "white") +
  geom_smooth(aes(color = party), method = "lm") +
  geom_point(aes(fill = party), shape = 21, color = "black") +
  scale_x_continuous(
    breaks = scales::breaks_width(10),
    labels = scales::label_percent(scale = 1)
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(0.1),
    labels = scales::label_percent(accuracy = 1)
  ) +
  scale_fill_manual(
    breaks = c("Republican", "Democrat"),
    values = c("red3", "royalblue"),
    aesthetics = c("color", "fill")
  ) +
  coord_fixed(ratio = 60) +
  theme(legend.position = "bottom") +
  labs(
    title = "Trump Score vs. State Net Trump Vote",
    subtitle = "States that favored Trump have senators that vote with Trump more often",
    caption = "Sources: FiveThirtyEight",
    x = "State Net Trump Vote (2016)",
    y = "Trump Score",
    color = NULL,
    fill = NULL
  ) +
  ggsave("figures/trump_score_vs_net_trump_vote.png")
```

```{r} 
member_averages_df %>% 
  filter(congress == 115L) %>% 
  ggplot(aes(votes)) +
  geom_dotplot(binwidth = 1) +
  labs(
    title = "Number of Votes by Senator, 115th Congress",
    x = "Number of Votes",
    y = ""
  ) +
  ggsave("figures/trump_score_votes.png")
```


## Trump Score over time

```{r}
trump_score <- 
  tibble(
    date = 
      seq(START_115, END_115, by = "1 day") %>% as.Date()
  ) %>% 
  left_join(
    votes_df %>% 
      mutate(date = as.Date(voted_at)) %>% 
      group_by(date, bioguide, last_name) %>% 
      summarize(
        agrees = sum(agree),
        votes = n()
      ) %>% 
      pivot_wider(
        names_from = c(bioguide, last_name),
        values_from = c(agrees, votes)
      ),
    by = "date"
  ) %>% 
  mutate_at(
    vars(starts_with("agrees"), starts_with("votes")),
    RcppRoll::roll_sum,
    n = window,
    align = "right",
    fill = NA_integer_,
    na.rm = TRUE
  ) %>% 
  slice(-(1:window)) %>% 
  pivot_longer(
    cols = -date,
    names_to = c("variable", "bioguide", "last_name"),
    names_sep = "_",
    values_to = "value"
  ) %>%
  pivot_wider(
    names_from = variable,
    values_from = value
  ) %>%
  filter(votes > 0) %>% 
  mutate(trump_score = agrees / votes) %>% 
  left_join(
    member_averages_df %>% 
      filter(congress == 115) %>% 
      select(bioguide, party),
    by = "bioguide"
  )

trump_score %>% 
  drop_na(trump_score) %>% 
  filter(party != "Independent") %>% 
  group_by(date, party) %>% 
  summarize(
    pct_25 = quantile(trump_score, 0.25),
    pct_50 = quantile(trump_score, 0.5),
    pct_75 = quantile(trump_score, 0.75)
  ) %>% 
  ggplot(aes(date, fill = party)) +
  geom_line(aes(y = pct_25, group = party)) +
  geom_line(aes(y = pct_75, group = party)) +
  geom_ribbon(aes(ymin = pct_25, ymax = pct_75, alpha = "middle-50")) +
  geom_line(aes(y = pct_50, group = party)) +
  geom_hline(yintercept = 0.5) +
  scale_x_date(
    date_breaks = "2 months",
    date_labels = "%b %y"
  ) +
  scale_y_continuous( 
    breaks = scales::breaks_width(0.1),
    labels = scales::label_percent(accuracy = 1)
  ) +
  scale_fill_manual(
    breaks = c("Republican", "Democrat"),
    values = c("red", "royalblue")
  ) +
  scale_alpha_manual(
    breaks = c("middle-50"),
    labels = c("Middle 50%"),
    values = c(0.85)
  ) +
  coord_cartesian(
    ylim = c(0, 1),
    expand = FALSE
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  ) +
  labs(
    title = str_glue("Middle 50% of {window}D Rolling Trump Scores by Party, 115th Congress"),
    x = "",
    y = "Trump Score",
    fill = "",
    alpha = ""
  ) +
  ggsave("figures/trump_score_party.png")
```

Deciding whether to use FiveThirtyEight's predicted Trump scores. They don't seem
very well calibrated, so I'm thinking I will either have to ditch the idea of controlling
the votes for expectation altogether or come up with my own way to do this.

```{r fig.asp=1}
votes_df %>% 
  transmute(
    bioguide = bioguide,
    last_name = last_name,
    state = state,
    party = party,
    prob_interval = cut_width(predicted_probability, width = 0.1, center = 0.05, label = FALSE),
    agree = if_else(agree == -1, NA_integer_, agree)
  ) %>% 
  group_by(prob_interval, party) %>% 
  summarize(
    n_sample = n(),
    agree_pct = mean(agree, na.rm = TRUE)
  ) %>% 
  filter(party != "Independent") %>% 
  ggplot(aes(prob_interval, agree_pct)) +
  geom_line(
    aes(group = 1),
    data = 
      tribble(
        ~ prob_interval, ~ agree_pct,
        0, 0.05,
        1, 0.95,
      ) %>% 
      mutate(prob_interval = cut_width(prob_interval, width = 0.1, center = 0.05, label = FALSE)),
    color = "red"
  ) +
  geom_point(aes(size = n_sample)) +
  geom_line(aes(group = 1)) +
  scale_x_continuous(
    breaks = 1:10,
    labels = c(
      "0% - 10%",
      "10% - 20%",
      "20% - 30%",
      "30% - 40%",
      "40% - 50%",
      "50% - 60%",
      "60% - 70%",
      "70% - 80%",
      "80% - 90%",
      "90% - 100%"
    )
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(0.2),
    labels = scales::label_percent(accuracy = 1)
  ) +
  scale_size_area(
    breaks = scales::breaks_log(n = 4)
  ) +
  facet_wrap(~ party) +
  theme(
    legend.position = "bottom",
    axis.text = element_text(size = 6)
  ) +
  labs(
    title = "Calibration Plot: FiveThirtyEight's Trump Agreement % Model",
    x = "Prediction Interval",
    y = "Observed Agreement %",
    size = "No. Obs."
  ) +
  ggsave("figures/predicted_trump_score_calibration.png")
```


