---
title: "Relating Protest Attendance and Senator Trump Score"
author: "Mac Bagwell"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(lubridate)
library(modelr)

# Parameters
WINDOW <- 90
LAMBDA <- 0.95

protest_file <- here::here("c01-own/data/protests_population.rds")
member_averages_file <- here::here("c01-own/data/member_averages.rds")
votes_file <- here::here("c01-own/data/votes.rds")
class_file <- here::here("c01-own/data/senate-class.rds")
#===============================================================================

# Code
ewma <- function(x, lambda, init = median(x, na.rm = TRUE)) {
	p <- init
  if_else(!is.finite(x), 0, x) %>% 
	  map_dbl(~ (p <<- p * lambda + (.) * (1 - lambda)))
}

protest_trump_plot <- function(state_) {
  state_plot <- 
    cowplot::plot_grid(
      state_protests_ts %>% 
        filter(state == state_) %>% 
        ggplot(aes(date, protest_attendance)) +
        geom_line() +
        scale_x_date(date_breaks = "3 months") +
        scale_y_log10() +
        labs(
          title = 
            str_c(
              state_, 
              str_glue("EWMA Protest Attendance"), 
              sep = " "
            ),
          x = "Date",
          y = "Protest Attendance per 100,000 People"
        ),
      trump_score_ts %>% 
        filter(state == state_) %>% 
        ggplot(aes(date, trump_score_forward, color = last_name)) +
        geom_line() +
        scale_x_date(date_breaks = "3 months") +
        scale_y_continuous(breaks = seq(0, 1, by = 0.25)) +
        theme(legend.position = "bottom") +
        labs(
          title = str_c(state_, str_glue("Senator Trump Score, Next {WINDOW} Days"), sep = " "),
          x = "Date",
          y = "Trump Score",
          color = "Senator"
        ),
      nrow = 2
    )
  
  state_plot +
    ggsave(
      str_glue("c01-own/eda/figures/{state_}_ts.png") %>% 
        here::here()
    )
  
  state_plot
}

protest_trump_scatter <- function(state_) {
  state_protests_ts %>% 
    filter(state == state_) %>% 
    left_join(trump_score_ts %>% filter(state == state_), by = c("date", "state")) %>% 
    drop_na(last_name, protest_attendance, contains("trump_score")) %>% 
    mutate(trump_pm = trump_score_diff) %>% 
    ggplot(aes(protest_attendance, trump_pm, fill = date)) +
    geom_hline(yintercept = 0, size = 1.5, color = "white") +
    geom_smooth(method = "lm", color = "red", se = FALSE) +
    geom_jitter(shape = 21, color = "white", stroke = 0.1) +
    scale_x_log10() +
    facet_wrap(vars(last_name)) +
    theme(legend.position = "bottom") +
    labs(
      title = 
        str_c(
          state_, 
          str_glue("Senator Change in {WINDOW}-day Plus-Minus vs. Protest Attendance"), 
          sep = " "
        ),
      caption = "Source: FiveThirtyEight, Crowd Counting Consortium",
      x = "EWMA Protest Attendance",
      y = "Trump PM"
    ) +
    ggsave(
      str_glue("c01-own/eda/figures/{state_}_scatter.png") %>% 
        here::here()
    ) 
}

protest_population <- read_rds(protest_file) %>% filter(pro_anti_trump == "anti")
member_averages <- 
  read_rds(member_averages_file) %>% 
  filter(congress == 115L) %>% 
  left_join(read_rds(class_file), by = c("last_name", "state"))
votes <- read_rds(votes_file) 

state_protests_ts <- 
  tibble(
    date = 
      seq(ymd("2017-02-02"), ymd("2018-12-31"), by = "1 day") %>% as.Date()
  ) %>% 
  left_join(
    protest_population %>% 
      count(state, date = as.Date(date), wt =  attendance / (population / 1e5)) %>% 
      pivot_wider(
        names_from = state,
        values_from = n,
        values_fill = list(n = 0L)
      ),
    by = "date"
  ) %>% 
  mutate_at(
    vars(-date),
    ewma,
    lambda = LAMBDA
  ) %>%
  pivot_longer(
    cols = -date,
    names_to = "state",
    values_to = "protest_attendance"
  )

trump_score_ts <- 
  tibble(
    date = 
      seq(ymd("2017-02-02"), ymd("2018-12-31"), by = "1 day") %>% as.Date()
  ) %>% 
  left_join(
    votes %>% 
      semi_join(
        member_averages %>% 
          filter(congress == 115L, votes >= 80, party != "Independent"),
        by = "bioguide"
      ) %>% 
      mutate(date = as.Date(voted_at)) %>% 
      group_by(date, bioguide, last_name, state, party) %>% 
      summarize(
        agrees = sum(agree),
        predicted = sum(predicted_probability),
        votes = n()
      ) %>% 
      pivot_wider(
        names_from = c(bioguide, last_name, state, party),
        values_from = c(agrees, predicted, votes)
      ),
    by = "date"
  ) %>% 
  replace_na(list(agrees = 0L, predicted = 0L, votes = 0L)) %>% 
  mutate_at(
    vars(
      starts_with("agrees"), 
      starts_with("predicted"), 
      starts_with("votes")
    ),
    list(
      forward = 
        ~ RcppRoll::roll_suml(., n = WINDOW, fill = NA_integer_, na.rm = TRUE),
      past = 
        ~ lag(.) %>% 
          RcppRoll::roll_sumr(., n = WINDOW, fill = NA_integer_, na.rm = TRUE)
    )
  ) %>% 
  select(date, ends_with("forward"), ends_with("past")) %>% 
  pivot_longer(
    cols = -date,
    names_to = c("variable", "bioguide", "last_name", "state", "party", "window"),
    names_sep = "_",
    values_to = "value"
  ) %>% 
  pivot_wider(
    names_from = c(variable, window),
    values_from = value
  ) %>%
  mutate(
    trump_score_forward = (agrees_forward + 1) / (votes_forward + 2),
    predicted_trump_score_forward = predicted_forward / votes_forward,
    trump_score_past = (agrees_past + 1) / (votes_past + 1),
    predicted_trump_score_past = predicted_past / votes_past,
    trump_score_diff = trump_score_forward - trump_score_past,
    predicted_trump_score_diff = 
      predicted_trump_score_forward - predicted_trump_score_past,
    last_name = 
      str_c(
        last_name,
        case_when(
          party == "Republican" ~ str_glue("(R-{state})"),
          party == "Democrat" ~ str_glue("(D-{state})"),
          party == "Independent" ~ str_glue("(I-{state})"),
          TRUE ~ "(Other)"
        ),
        sep = " "
      )
  )
```

## Recreating Miras' Plot

Below we recreate a slightly different version of Miras' plot. The
x-axis is the log of total anti-Trump protest attendance during the 115th session of congress
by senators' state scaled by population, and the y-axis is the "Trump Score" of
the senator. We see a negative correlation between the two for the Democrats,
and a weaker such trend for Republicans. But this could easily be attributed
to the relative political tastes of the constituencies, and protests need not
have any causal relationship with Senatorial legilative activity.

```{r}
protest_population %>% 
  count(
    state, 
    name = "protest_activity", 
    wt = attendance / (population / 1e5)
  ) %>% 
  left_join(
    member_averages %>%
      select(state, party, agree_pct, votes) %>% 
      filter(party != "Independent"),
    by = "state"
  ) %>% 
  ggplot(aes(protest_activity, agree_pct, fill = party)) +
  geom_smooth(aes(color = party), method = "lm", formula = y ~ log(x)) +
  geom_point(shape = 21, color = "black") +
  scale_color_manual(
    breaks = c("Republican", "Democrat"),
    values = c(Republican = "red3", Democrat = "royalblue"),
    aesthetics = c("color", "fill")
  ) +
  scale_x_log10(
    breaks = scales::breaks_log(n = 10)
  ) +
  scale_y_continuous(
    labels = scales::label_percent()
  ) +
  coord_fixed(ratio = 1.5) +
  theme(legend.position = "bottom") +
  labs(
    title = "Senator Trump Score vs. State Protest Activity, 115th Congress",
    subtitle = "Independents omitted",
    caption = "Sources: FiveThirtyEight, Crowd Counting Consortium",
    x = "Protest Attendance per 100,000 People, Feb. 2017 - Dec. 2018",
    y = "Trump Score",
    color = NULL,
    fill = NULL
  ) +
  ggsave("figures/miras.png")
```

## Correlation Between Protest Activity and Trump's Performance

As suspected, Trump's performance is somewhat correlated with protest activity
in that state. We ought to control for this before considering the effect
of protests on senators.

```{r}
protest_trump_state <- 
  protest_population %>% 
  count(
    state, 
    name = "protest_activity", 
    wt = attendance / (population / 1e5)
  ) %>% 
  left_join(
    member_averages %>%
      filter(congress == 115L, party != "Independent") %>% 
      select(state, net_trump_vote, party, agree_pct),
    by = "state"
  )

protest_trump_state %>% 
  ggplot(aes(net_trump_vote, protest_activity)) +
  geom_vline(xintercept = 0, color = "white", size = 1.5) +
  geom_smooth(method = "lm") +
  geom_text(aes(label = state)) +
  scale_x_continuous(
    labels = scales::label_percent(scale = 1)
  ) +
  scale_y_log10(
    breaks = scales::breaks_log(n = 9)
  ) +
  labs(
    title = "Protest Attendance per 100,000 people vs. Net Trump Vote",
    subtitle = "States that favored Trump are less active",
    caption = "Sources: FiveThirtyEight, Crowd Counting Consortium",
    x = "Trump's Margin of Victory, 2016",
    y = "Protest Attendance per 100,000 people, Feb. 2017 - Dec. 2018"
  ) +
  ggsave("figures/protest_vs_net_trump_vote.png")
```

```{r}
protest_trump_mod <- 
  protest_trump_state %>% 
  fit_with(
    lm, 
    list(
      protest = log(protest_activity) ~ net_trump_vote,
      trump = agree_pct ~ net_trump_vote * party
    )
  )

protest_population %>% 
  count(
    state, 
    name = "protest_activity", 
    wt = attendance / (population / 1e5)
  ) %>% 
  left_join(
    member_averages %>%
      select(state, party, agree_pct, net_trump_vote, votes) %>% 
      filter(party != "Independent", votes >= 80),
    by = "state"
  ) %>% 
  spread_residuals(!!! protest_trump_mod) %>% 
  ggplot(aes(protest, trump, fill = party)) +
  geom_hline(yintercept = 0, size = 1.5, color = "white") +
  geom_vline(xintercept = 0, size = 1.5, color = "white") +
  geom_point(shape = 21, size = 2, color = "black") +
  scale_fill_manual(
    breaks = c("Republican", "Democrat"),
    values = c(Republican = "red3", Democrat = "royalblue")
  ) +
  scale_y_continuous(
    labels = scales::label_percent()
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = "Residual Trump Score vs. Residual Protest Activity during the 115th Congress",
    subtitle = "Distribution does not suggest residual effects",
    caption = "Variables controlled for Trump's 2016 margin of victory by state\nSources: FiveThirtyEight, Crowd Counting Consortium",
    x = "Residual Protest Activity, Feb. 2017 - Dec. 2018",
    y = "Residual Trump Score",
    fill = NULL
  ) +
  ggsave("figures/miras-resid.png")
```


## Correlation Between Protest Attendance and Trump Score

Preliminary looks at the time series relationship between state's protest attendance
and senators' Trump Scores don't reveal any patterns, contrary to Miras' findngs. We would
expect to find that Trump Score goes down as protest activity increases, However,
looking at some of the time series plots does not suggest compelling relationships


```{r eval=FALSE}
map(state.abb, protest_trump_plot) %>% 
  walk(print)
```

```{r eval=FALSE}
map(state.abb, protest_trump_scatter) %>% 
  walk(print)
```


```{r}
corrs <- 
  state_protests_ts %>% 
  left_join(
    trump_score_ts %>% 
      semi_join(
        member_averages %>% 
          filter(congress == 115, votes > 80),
        by = "bioguide"
      ), 
    by = c("date", "state")
  ) %>% 
  filter(date >= "2017-02-16", party != "Independent") %>% 
  group_by(bioguide, last_name, party, state) %>% 
  summarize(
    r = 
      cor(
        log(protest_attendance), 
        trump_score_forward - predicted_trump_score_forward, 
        method = "spearman", 
        use = "pair"
      )
  ) %>% 
  ungroup() %>% 
  left_join(
    member_averages %>% 
      select(bioguide, class, net_trump_vote), 
    by = "bioguide"
  )

corrs %>% 
  ggplot(aes(r, fill = party)) +
  geom_vline(xintercept = 0, color = "white", size = 1.5) +
  geom_dotplot(alpha = 0.8, binwidth = 0.03) +
  scale_x_continuous(
    breaks = scales::breaks_width(0.2)
  ) +
  scale_fill_manual(
    breaks = c("Democrat", "Republican"),
    values = c("royalblue", "red3")
  ) + 
  theme(
    legend.position = "bottom",
    axis.ticks.length.y = unit(0, "mm"),
    axis.text.y = element_blank()
  ) +
  labs(
    title = str_glue("Correlation between {WINDOW}D Forward Trump Score and Protest Activity"),
    subtitle = "Distribution does not suggest negative effects",
    caption = "Spearman method used to compute correlation\nSources: FiveThirtyEight, Crowd Counting Consortium",
    x = "Correlation Coefficient",
    y = NULL,
    fill = NULL
  ) +
  ggsave("figures/trump_score_protest_spearman.png")
```

```{r}
corrs %>% 
  ggplot(aes(r, fill = party)) +
  geom_vline(xintercept = 0) +
  geom_dotplot(binwidth = 0.025, alpha = 0.8) +
  scale_x_continuous(
    breaks = scales::breaks_width(0.1)
  ) +
  scale_fill_manual(
    breaks = c("Democrat", "Republican"),
    values = c(Democrat = "royalblue", Republican = "red3")
  ) + 
  facet_grid(
    rows = vars(class), 
    labeller = labeller(class = function(x) str_c("Class ", x))
  ) +
  coord_fixed(ratio = 0.25) +
  theme(
    plot.title = element_text(size = 10),
    plot.subtitle = element_text(size = 8),
    plot.caption = element_text(size = 6),
    strip.text = element_text(size = 8),
    axis.title = element_text(size = 8),
    axis.text = element_text(size = 6),
    axis.text.y = element_blank(),
    axis.ticks.length.y = unit(0, "mm"),
    legend.position = "none"
  ) +
  labs(
    title = str_glue("Correlation between change in {WINDOW}-Day Trump Score and Protest Activity"),
    subtitle = "Class 1 up for re-election in 2018, Class 2 in 2020, Class 3 in 2022",
    caption = "Spearman method used to compute correlation\nSources: FiveThirtyEight, Crowd Counting Consortium, Wikipedia",
    x = "Correlation Coefficient",
    y = NULL
  ) +
  ggsave("figures/trump_protest_spearman_class_dot.png")

corrs %>% 
  ggplot(aes(factor(class), r, fill = party)) +
  geom_hline(yintercept = 0) +
  geom_boxplot(alpha = 0.8) +
  scale_y_continuous(
    breaks = scales::breaks_width(0.1)
  ) +
  scale_fill_manual(
    breaks = c("Democrat", "Republican"),
    values = c(Democrat = "royalblue", Republican = "red3")
  ) +
  coord_flip() +
  theme(
    plot.title = element_text(size = 12),
    plot.subtitle = element_text(size = 10),
    plot.caption = element_text(size = 8),
    strip.text = element_text(size = 8),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8),
    legend.position = "bottom"
  ) +
  labs(
    title = str_glue("Correlation between change in {WINDOW}-Day Trump Score and Protest Activity"),
    subtitle = "Class 1 up for re-election in 2018, Class 2 in 2020, Class 3 in 2022",
    caption = "Spearman method used to compute correlation\nSources: FiveThirtyEight, Crowd Counting Consortium, Wikipedia",
    x = "Class",
    y = "Correlation Coefficient",
    fill = NULL
  ) +
  ggsave("figures/trump_protest_spearman_class_box.png")
```

```{r fig.asp=1}
corrs %>% 
  left_join(
    protest_population %>% 
      count(
        state, 
        name = "protest_activity", 
        wt = attendance / (population / 1e5)
      ),
    by = "state"
  ) %>% 
  ggplot(aes(net_trump_vote, r)) +
  geom_hline(yintercept = 0, size = 1.5, color = "white") +
  geom_vline(xintercept = 0, size = 1.5, color = "white") +
  geom_point(aes(fill = party), shape = 21, size = 2) +
  ggrepel::geom_text_repel(
    aes(label = last_name, color = party),
    size = 2
  ) +
  scale_x_continuous(
    breaks = scales::breaks_width(20),
    labels = scales::label_percent(scale = 1, accuracy = 1)
  ) +
  scale_color_manual(
    breaks = c("Democrat", "Republican"),
    values = c(Democrat = "royalblue", Republican = "red3"),
    aesthetics = c("color", "fill")
  ) +
  coord_fixed(ratio = 40) +
  facet_grid(
    rows = vars(class),
    labeller = labeller(class = function(x) str_c("Class ", x))
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = str_glue("Correlation coefficient vs. Trump Victory Margin in State"),
    subtitle = "Class 1 up for re-election in 2018, Class 2 in 2020, Class 3 in 2022",
    caption = "Spearman method used to compute correlation\nSources: FiveThirtyEight, Crowd Counting Consortium, Wikipedia",
    x = "Trump Victory Margin",
    y = "Correlation Coefficient",
    color = NULL,
    fill = NULL
  ) +
  ggsave("figures/trump_protest_corr_victory_margin_class.png", width = 10, height = 14)
```

```{r}
corrs %>% 
  left_join(
    protest_population %>% 
      count(
        state, 
        name = "protest_activity", 
        wt = attendance / (population / 1e5)
      ),
    by = "state"
  ) %>% 
  ggplot(aes(protest_activity, r)) +
  geom_hline(yintercept = 0, size = 1.5, color = "white") +
  geom_point(aes(fill = party), shape = 21, size = 2) +
  ggrepel::geom_text_repel(
    aes(label = last_name, color = party),
    size = 2
  ) +
  scale_x_log10(
    breaks = scales::breaks_log(n = 10)
  ) +
  scale_color_manual(
    breaks = c("Democrat", "Republican"),
    values = c(Democrat = "royalblue", Republican = "red3"),
    aesthetics = c("color", "fill")
  ) +
  coord_fixed(ratio = 0.5) +
  facet_grid(
    rows = vars(class),
    labeller = labeller(class = function(x) str_c("Class ", x))
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = str_glue("Correlation coefficient vs. Protest Attendance in State"),
    subtitle = "Class 1 up for re-election in 2018, Class 2 in 2020, Class 3 in 2022",
    caption = "Spearman method used to compute correlation\nSources: FiveThirtyEight, Crowd Counting Consortium, Wikipedia",
    x = "Total Protest Attendance per 100,000",
    y = "Correlation Coefficient",
    color = NULL,
    fill = NULL
  ) +
  ggsave("figures/trump_protest_corr_protest_activity_class.png", width = 10, height = 14)
```


