---
title: "Population-Adjusted Protest Data"
author: "Mac Bagwell"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(lubridate)

# Parameters
protests_population_file <- here::here("c01-own/data/protests_population.rds")

#===============================================================================

# Code
protests_population <- read_rds(protests_population_file)
```

## Population Adjusted Protest Attendance 2017-18

As we did before, we reduce the estimates data to one point estimate
depending on whether a point estimate is provided or a range is provided.


```{r}
protests_population %>% 
  filter(pro_anti_trump == "anti") %>% 
  count(state, population, wt = attendance) %>%  
  ggplot(aes(population, n)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_log10() +
  scale_y_log10() +
  labs(
    title = "Protest Attendance vs. State Population, Feb 2017 - Dec 2018",
    x = "State Population",
    y = "Protest Attendance"
  ) +
  ggsave("figures/protest_vs_population.png")
  
```

Proportional to population, New York was actually the most
active Anti-Trump protest state, and California was actually less active than 
states like Conneticut, Massachussetts, and Washington. This differs from the perspective
given by Miras' measure of "large protests".

There is still significant variation in protest activity against Trump that isn't
explained by differing populations. I would think that Trump's performance in each
state in 2016 probably explains a lot of this variation.

```{r}
protests_population %>% 
  count(state, wt = attendance / (population / 1e5)) %>% 
  mutate(state = state %>% fct_reorder(n)) %>% 
  ggplot(aes(state, n)) +
  geom_point() +
  scale_x_discrete(
    breaks = state.abb,
    labels = set_names(state.name, state.abb)
  ) +
  scale_y_log10(
    breaks = scales::breaks_log(n = 10)
  ) +
  coord_flip() +
  labs(
    title = "Protest Attendance Per Population (Anti-Trump only)",
    x = NULL,
    y = "Protest Attendance per 100,000 people"
  ) +
  ggsave("figures/protest_per_population_by_state.png")
```

```{r}
ewma <- function(x, lambda) {
	p <- 0
  if_else(!is.finite(x), 0, x) %>% 
	  map_dbl(
	    ~ (p <<- p * lambda + . * (1 - lambda))
	   )
}

protests_population %>% 
  filter(month(date) %in% 7:9, year(date) == 2018) %>% 
  arrange(date)
  count(
    date = floor_date(date, unit = "day"), 
    wt = attendance,
    name = "attendance"
  ) %>% 
  mutate(
    ewma_attendance = ewma(attendance, lambda = 0.95)
  ) %>% 
  ggplot(aes(date)) +
  geom_segment(aes(xend = date, y = 1, yend = 1 + attendance), alpha = 0.5) +
  geom_line(aes(y = 1 + ewma_attendance), color = "red") +
  scale_x_date(
    date_breaks = "3 months"
  ) +
  scale_y_log10() + 
  labs(
    title = "National Protest Attendance over Time",
    subtitle = "Red line represents EMA of daily protest attendance per 100,000 people",
    caption = "Source: Crowd Counting Consortium",
    x = "Date",
    y = "National Protest Attendance per 100,000"
  ) +
  ggsave("figures/protest_vs_time.png")
```


